---
title: '**Visualisation**: Assignment 1'
author: 'Name: Roudranil Das Roll Number: MDS202227'
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instruction: 
1) Work on the `Assignment1.Rmd' file. Compile the file as pdf. Submit only the pdf file in moodle.
2) Make sure you write your name and roll number at the top of the `Assignment1.Rmd' file.

## Total Marks: 10 points

## Problem 1 (3 points)

**Problem Statement:** Write an $\texttt{R}$ function which will test Central Limit Theorem.


*   Assume the underlying population distribution follow Poisson distribution with rate parameter $\lambda$
*   We want to estimate the unknown $\lambda$ with the sample mean 
$$
\hat{\lambda}=\frac{1}{n}\sum_{i=1}^{n}X_i
$$
*   The exact sampling distribution of $\hat{\lambda}$ is unknown
*   But CLT tells us that as sample size $n$ increases the sampling distribution of $\hat{\lambda}$ can be approximated by Gaussian distribution.


### **Input in the function:**

* $n$: sample size 
* $\lambda$ : rate parameter
* $N$: simulation size


## **Output from the function:** 

* Histogram of the sampling distribution using `ggplot`
* QQ-plot using `ggplot`

## **Test cases:**

* case 1 a: $\lambda=0.7$, n=10, N=5000
* case 1 b: $\lambda=0.7$, n=30, N=5000
* case 1 c: $\lambda=0.7$, n=100, N=5000
* case 1 d: $\lambda=0.7$, n=300, N=5000

* case 2 a: $\lambda=1.7$, n=10, N=5000
* case 2 b: $\lambda=1.7$, n=30, N=5000
* case 2 c: $\lambda=1.7$, n=100, N=5000
* case 2 d: $\lambda=1.7$, n=300, N=5000

```{r}
library(tidyverse)

simulate_clt = function(lambda, n, N){
    sim_means = c()
    for(i in 1:N){
        sample = rpois(n, lambda)
        sample_mean = mean(sample)
        sim_means = append(sim_means, sample_mean)
    }
    # we create a dataset with the columns being the sample means and 
    # respective values of lambda and n
    data = data.frame(sim_means, rep(lambda, N), rep(n, N))
    colnames(data) = c("mean", "lambda", "n")
    return(data)
}

lambdas = rep(c(0.7, 1.7), times = 1, each = 4)
n = rep(c(10, 30, 100, 300), times = 2)
N = rep(5000, times = 8)
```

For $\lambda = 0.7$, we get the following $4$ histograms:
```{r fig=TRUE, fig.dim=c(7,3)}
dataset = data.frame()
for(i in 1:4){
    # the datasets all created are binded rowwise
    # this is done to help us to create the facet grids
    dataset = rbind(dataset, simulate_clt(lambdas[i], n[i], N[i]))
}
dataset %>% 
    ggplot(aes(x = mean)) + 
    geom_histogram(bins = 20, color = "white") + 
    facet_grid(vars(lambda), vars(n), scales = "free") + 
    labs(title = paste("Histograms corresponding to lambda =", lambdas[1], sep = " "),
         x = "Sample means",
         y = "Count")
```

For the same value of $\lambda$, we get the following $4$ QQ-plots.
```{r fig.dim=c(5,5)}
dataset = data.frame()
for(i in 1:4){
    dataset = rbind(dataset, simulate_clt(lambdas[i], n[i], N[i]))
}
dataset %>% 
    ggplot(aes(sample = mean)) +
    geom_qq(size = 0.6) +
    geom_qq_line(color = "blue") +
    facet_wrap(vars(n), scales = "free") +
    labs(title = "QQ-plot for lambda = 0.7")
```

For $\lambda = 1.7$, we get the following 4 histograms:
```{r fig=TRUE, fig.dim=c(7,3)}
dataset = data.frame()
for(i in 5:8){
    dataset = rbind(dataset, simulate_clt(lambdas[i], n[i], N[i]))
}
dataset %>% 
    ggplot(aes(x = mean)) + 
    geom_histogram(bins = 20, color = "white") + 
    facet_grid(vars(lambda), vars(n), scales = "free") + 
    labs(title = paste("Histograms corresponding to lambda =", lambdas[5], sep = " "),
         x = "Sample means",
         y = "Count")
```

And we follow this the corresponding QQ-plots:
```{r fig.dim=c(5,5)}
dataset = data.frame()
for(i in 5:8){
    dataset = rbind(dataset, simulate_clt(lambdas[i], n[i], N[i]))
}
dataset %>% 
    ggplot(aes(sample = mean)) +
    geom_qq(size = 0.6) +
    geom_qq_line(color = "blue") +
    facet_wrap(vars(n), scales = "free") +
    labs(title = "QQ-plot for lambda = 1.7")
```

### **Problem 2:** (2 points)

Consider the `JohnsonJohnson` dataset. The datset contains the Quarterly earnings (dollars) per Johnson & Johnson share 1960–80.

a) Draw the time series plot of Quarterly earnings in regular scale and log-scale using the ` ggplot` (1 point)

```{r}
head(JohnsonJohnson)
```

```{r}
JohnsonJohnson %>% 
    ggplot(aes(x))
```


### **Problem 3:** (2 points)

- An experiment was conducted to measure and compare the effectiveness of various feed supplements on the growth rate of chickens.

- Following R-code is a standard side-by-side boxplot showing effect of feed supplements on the growth rate of chickens.

```{r,fig=T}
boxplot(weight~feed,data=chickwts,pch=20
        ,main = "chickwt data"
        ,ylab = "Weight at six weeks (gm)"
        ,xlab = "Feed Type")
```


a) Reproduce the same plot using the `ggplot`; while fill each boxes with different colour. 

b) In addition draw probability density plot for weights of chicken's growth by each feed seperately using the `ggplot`. Draw this plot seperately. 

```{r}
## write your R-function for problem 3 here
##
##
```

### **Problem 4:** (3 points)

Consider the `EuStockMarkets` data available in `R`. Contains the daily closing prices of major European stock indices: Germany DAX (Ibis), Switzerland SMI, France CAC, and UK FTSE. The data are sampled in business time, i.e., weekends and holidays are omitted.

```{r}
head(EuStockMarkets)
```

+ Suppose $P_t$ is the closing price of a stock indices on day $t$. 
+ The daily return $r_t$ is defined as
$$
r_t = \log(P_t)-\log(P_{t-1}).
$$
a) Draw time-series plot of $P_t$ for all four markets
b) Draw time-series plot of $r_t$ for all four markets
c) Draw histogram of $P_t$ for all four markets
d) Draw histogram of $r_t$ for all four markets
e) Suppose you invested \$ 1000 in each market indices on day 1. Plot how your investment grows on the same plot for all four markets. Make your plot using `ggplot`.
f) Check which market outperform others during the same time?

Make all your plots using `ggplot`.

```{r}
## write your R-function for problem 4 here
##
##
```
